---
title: 'The decennial Census with tidycensus'
teaching: 10
exercises: 2
---

:::::::::::::::::::::::::::::::::::::: questions 

- How do you import decennial Census data into R Studio using *tidycensus*?
- How do you process, query, wrangle, and visualize decennial Census data once it is imported into R Studio? 

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Explain how to use *tidycensus* to import decennial Census data at various geographic scales, and various temporal periods, into your R Studio environment. 
- Process and query decennial Census data to explore simple questions, using the *dplyr* package 
- Visualize decennial Census data using the *ggplot2* and *tmap packages
- Export decennial Census datasets and visualizations from your R Studio environment to a local directory in various file formats. 

::::::::::::::::::::::::::::::::::::::::::::::::

## Introduction

```{r, message=FALSE, warning=FALSE}
library(tidycensus)
library(tmap)
library(tidyverse)
```

```{r, echo=F, message=FALSE, warning=FALSE}
library(kableExtra)
library(here)
```

## Retrieving decennial Census data

```{r, eval=FALSE}
# loads variable table for 2010 decennial census and assigns it to object named
# "decennial_variables_2010"
 decennial_variables_2010<-load_variables(2010, "sf1")
```

```{r, echo=FALSE, message=FALSE}
decennial_variables_2010 <- read_csv("data/decennial_variables_2010.csv")
```

```{r}
# prints "decennial_variables_2010"
decennial_variables_2010
```

```{r, eval=F}
# brings up "decennial_variables_2010" in data Viewer
View(decennial_variables_2010)
```


```{r, echo=FALSE}
decennial_variables_2010 %>%
  head(40) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  row_spec(0, background = "#333", color = "white") %>%  # Fix header
  scroll_box(height = "300px", width = "100%")
```


test

## Cleaning and processing decennial Census datasets

## Visualizing decennial Census data with *ggplot2*

## Mapping decennial Census data with *tmap*

## Working with multiple Census datasets

## Exporting decennial Census datasets and analysis products from R Studio


This is a lesson created via The Carpentries Workbench. It is written in
[Pandoc-flavored Markdown][pandoc] for static files (with extension `.md`) and
[R Markdown][r-markdown] for dynamic files that can render code into output
(with extension `.Rmd`). Please refer to the [Introduction to The Carpentries
Workbench][carpentries-workbench] for full documentation.

What you need to know is that there are three sections required for a valid
Carpentries lesson template:

 1. `questions` are displayed at the beginning of the episode to prime the
    learner for the content.
 2. `objectives` are the learning objectives for an episode displayed with
    the questions.
 3. `keypoints` are displayed at the end of the episode to reinforce the
    objectives.

:::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::: instructor

Inline instructor notes can help inform instructors of timing challenges
associated with the lessons. They appear in the "Instructor View"

::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: challenge 

## Challenge 1: Can you do it?

What is the output of this command?

```r
paste("This", "new", "lesson", "looks", "good")
```

:::::::::::::::::::::::: solution 

## Output
 
```output
[1] "This new lesson looks good"
```

:::::::::::::::::::::::::::::::::


## Challenge 2: how do you nest solutions within challenge blocks?

:::::::::::::::::::::::: solution 

You can add a line with at least three colons and a `solution` tag.

:::::::::::::::::::::::::::::::::
::::::::::::::::::::::::::::::::::::::::::::::::

## Figures

You can include figures generated from R Markdown:

```{r pyramid, fig.alt = "pie chart illusion of a pyramid", fig.cap = "Sun arise each and every morning"}
pie(
  c(Sky = 78, "Sunny side of pyramid" = 17, "Shady side of pyramid" = 5), 
  init.angle = 315, 
  col = c("deepskyblue", "yellow", "yellow3"), 
  border = FALSE
)
```
Or you can use pandoc markdown for static figures with the following syntax:

`![optional caption that appears below the figure](figure url){alt='alt text for
accessibility purposes'}`

![You belong in The Carpentries!](https://raw.githubusercontent.com/carpentries/logo/master/Badge_Carpentries.svg){alt='Blue Carpentries hex person logo with no text.'}

## Math

One of our episodes contains $\LaTeX$ equations when describing how to create
dynamic reports with {knitr}, so we now use mathjax to describe this:

`$\alpha = \dfrac{1}{(1 - \beta)^2}$` becomes: $\alpha = \dfrac{1}{(1 - \beta)^2}$

Cool, right?

::::::::::::::::::::::::::::::::::::: keypoints 

- Use `.md` files for episodes when you want static content
- Use `.Rmd` files for episodes when you need to generate output
- Run `sandpaper::check_lesson()` to identify any issues with your lesson
- Run `sandpaper::build_lesson()` to preview your lesson locally

::::::::::::::::::::::::::::::::::::::::::::::::

